---
title: "Enichment-analysis-fgsea"
author: "unawaz1996"
date: "2023-03-16"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = FALSE,
	warning = FALSE,
	message = FALSE
)
```

# Fast gene set enrichment analysis 
```{r load libraries}
source("code/libraries.R")
source("code/functions.R")
```

```{r}
theme_set(theme_bw())
panderOptions("table.split.table", Inf)
panderOptions("table.style", "rmarkdown")
panderOptions("big.mark", ",")
```

```{r load data}
load("output/limma-matrices.Rda")
load("output/DEG-limma-results.Rda")
load("output/DEG-list.Rda")
```

# Datasets



## Hallmark Gene sets 

```{r}
hm = msigdbr("Mus musculus", category = "H") %>% 
  distinct(gs_name, ensembl_gene, .keep_all = TRUE)
```

```{r}
c2 =  msigdbr("Mus musculus", category = "C2") %>% 
  dplyr::filter(gs_subcat == "CP:WIKIPATHWAYS" |
                          gs_subcat == "CP:KEGG"  | 
                          gs_subcat == "CP:REACTOME") %>% 
  distinct(gs_name, ensembl_gene, .keep_all = TRUE)

```


```{r}
load("/home/neuro/Documents/NMD_analysis/Analysis/Data/MSigDB/mouse/genesets_msigdbr.Rda")
```


Fast Gene Set Enrichment Analysis implements a fast version of the gsea algorithm. As a result, more permutations are made to get more fine grained p-values. 


`fgsea` results can be intrepreted as following: 

- `ES`: Enrichment score which is calculated based on the rank
- `NES`: Normalized Enrichment Score 


Using the MSigDB genesets as above, we will conduct a fgsea analysis on the sets of interest including KEGG, Wikipathways and Reactome. 


```{r}
fgsea_sets = all_gene_sets %>%
    dplyr::filter(gs_subcat == "CP:WIKIPATHWAYS" |
                          gs_subcat == "CP:KEGG"  | 
                          gs_subcat == "CP:REACTOME")

```

```{r include = FALSE}
fgsea_res = run_fgsea(v = v,  contrasts = contrasts, design = design, idx = fgsea_sets)
```


```{r}
sigRes = list()

sigRes[["UPF3B_KD_vs_Control"]] = fgsea_res$UPF3B_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes[["UPF3A_KD_vs_Control"]] = fgsea_res$UPF3A_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes[["UPF3A_OE_vs_Control"]] = fgsea_res$UPF3A_OE_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes[["DoubleKD_vs_Control"]] = fgsea_res$DoubleKD_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_res$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 
```


# UPF3B vs Knockdown results 

```{r}
fgsea_res$UPF3B_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="GSEA results for UPF3B KD compared to controls",
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```

```{r}
gsSizes = bind_rows(hm, c2) %>% 
  dplyr::select(gs_name, gene_symbol, ensembl_gene) %>% 
  chop(c(gene_symbol, ensembl_gene)) %>%
  mutate(
    gs_size = vapply(gene_symbol, length, integer(1)),
    de_id = lapply(
      X = ensembl_gene, 
      FUN = intersect, 
      y = dplyr::filter(DEGenes$UPF3B_vs_Control, Res.UPF3B_vs_Control != 0)$ensembl_gene_id
      ),
    de_size = vapply(de_id, length, integer(1))
  )

```

```{r fig.height=15, fig.width=10, fig.cap="*UpSet plot indicating distribution of DE genes within all significant gene sets. Gene sets were restricted to those with an FDR < 0.05 and at least 5 DE genes*"}
sigRes$UPF3B_KD_vs_Control %>% 
  dplyr::rename("gs_name" = fgsea.Geneset) %>%
  left_join(gsSizes) %>%
  dplyr::filter(de_size >= 5, fgsea.padj  < 0.05) %>%
  dplyr::select(gs_name, de_id) %>%
  unnest(de_id) %>% 
  split(f = .$gs_name) %>% 
  lapply(magrittr::extract2, "de_id") %>% 
  fromList() %>% 
  upset(
    nsets = length(.), 
    nintersects = 20,
    order.by = "freq", 
    mb.ratio = c(0.6, 0.4),
    sets.x.label = "Number Of DE Genes"
    )
```


```{r}
#upset(fromList(enrich_GSEA),  order.by = "freq")
```

# Gene sets enrichment in multiple comparisons 
```{r}
enrich_GSEA = list("UPF3B KD" = sigRes$UPF3B_KD_vs_Control$fgsea.Geneset, 
                "UPF3A KD" = sigRes$UPF3A_KD_vs_Control$fgsea.Geneset, 
                "Double KD" = sigRes$DoubleKD_vs_Control$fgsea.Geneset, 
                "UPF3A OE" = sigRes$UPF3A_OE_vs_Control$fgsea.Geneset, 
                "UPF3A OE, UPF3B KD" = sigRes$UPF3A_OE_UPF3B_KD_vs_Control$fgsea.Geneset)
```


## Hallmark sets 

```{r}
fgsea_sets = all_gene_sets %>%
    dplyr::filter(gs_cat == "H")
fgsea_resH = run_fgsea(v = v,  contrasts = contrasts, design = design, idx = fgsea_sets)

```

```{r}
sigRes_H = list()

sigRes_H[["UPF3B_KD_vs_Control"]] = fgsea_resH$UPF3B_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes_H[["UPF3A_KD_vs_Control"]] = fgsea_resH$UPF3A_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes_H[["UPF3A_OE_vs_Control"]] = fgsea_resH$UPF3A_OE_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes_H[["DoubleKD_vs_Control"]] = fgsea_resH$DoubleKD_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 

sigRes_H[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_resH$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
    dplyr::filter(fgsea.padj < 0.05) 
```

## UPF3B vs Controls 

```{r}
bind_rows(hm, c2) %>% 
  dplyr::select(gs_name, gene_symbol, ensembl_gene) %>% 
  chop(c(gene_symbol, ensembl_gene)) %>%
  mutate(
    gs_size = vapply(gene_symbol, length, integer(1)),
    de_id = lapply(
      X = ensembl_gene, 
      FUN = intersect, 
      y = dplyr::filter(DEGenes$UPF3B_vs_Control, Res.UPF3B_vs_Control != 0)$ensembl_gene_id
      ),de_size = vapply(de_id, length, integer(1))) %>% 
  right_join(dplyr::rename(sigRes_H$UPF3B_KD_vs_Contro, "gs_name" = fgsea.Geneset)) %>% 
  dplyr::filter(fgsea.padj  < 0.05) %>%
  dplyr::select(gs_name, de_id) %>%
  unnest(de_id) %>%
   mutate(gs_name = str_remove(gs_name, "HALLMARK_")) %>%
  split(f = .$gs_name) %>% 
  lapply(magrittr::extract2, "de_id") %>% 
  fromList() %>% 
  upset(
    nsets = length(.), 
    nintersects = 20,
    order.by = "freq", 
    mb.ratio = c(0.6, 0.4),
    sets.x.label = "Number Of DE Genes"
    )

```



## UPF3A vs Controls 

```{r}
bind_rows(hm, c2) %>% 
  dplyr::select(gs_name, gene_symbol, ensembl_gene) %>% 
  chop(c(gene_symbol, ensembl_gene)) %>%
  mutate(
    gs_size = vapply(gene_symbol, length, integer(1)),
    de_id = lapply(
      X = ensembl_gene, 
      FUN = intersect, 
      y = dplyr::filter(limma_results, Res.UPF3A_vs_Control != 0)$ensembl_gene_id
      ),
    de_size = vapply(de_id, length, integer(1))
  ) %>% right_join( dplyr::rename(sigRes_H$UPF3A_KD_vs_Control, "gs_name" = fgsea.Geneset)) %>% 
  dplyr::filter(fgsea.padj  < 0.05) %>%
  dplyr::select(gs_name, de_id) %>%
  unnest(de_id) %>%
   mutate(gs_name = str_remove(gs_name, "HALLMARK_")) %>%
  split(f = .$gs_name) %>% 
  lapply(magrittr::extract2, "de_id") %>% 
  fromList() %>% 
  upset(
    nsets = length(.), 
    nintersects = 20,
    order.by = "freq", 
    mb.ratio = c(0.6, 0.4),
    sets.x.label = "Number Of DE Genes"
    )

```
## Double KD vs Controls 

```{r}
bind_rows(hm, c2) %>% 
  dplyr::select(gs_name, gene_symbol, ensembl_gene) %>% 
  chop(c(gene_symbol, ensembl_gene)) %>%
  mutate(
    gs_size = vapply(gene_symbol, length, integer(1)),
    de_id = lapply(
      X = ensembl_gene, 
      FUN = intersect, 
      y = dplyr::filter(DEGenes$DoubleKD_vs_Control, Res.DoubleKD_vs_Control != 0)$ensembl_gene_id
      ),
    de_size = vapply(de_id, length, integer(1))
  ) %>% right_join( dplyr::rename(sigRes_H$DoubleKD_vs_Control, "gs_name" = fgsea.Geneset)) %>% 
  dplyr::filter(fgsea.padj  < 0.05) %>%
  dplyr::select(gs_name, de_id) %>%
  unnest(de_id) %>%
   mutate(gs_name = str_remove(gs_name, "HALLMARK_")) %>%
  split(f = .$gs_name) %>% 
  lapply(magrittr::extract2, "de_id") %>% 
  fromList() %>% 
  upset(
    nsets = length(.), 
    nintersects = 20,
    order.by = "freq", 
    mb.ratio = c(0.6, 0.4),
    sets.x.label = "Number Of DE Genes"
    )
```

## UPF3A OE vs Controls 

```{r}
bind_rows(hm, c2) %>% 
  dplyr::select(gs_name, gene_symbol, ensembl_gene) %>% 
  chop(c(gene_symbol, ensembl_gene)) %>%
  mutate(
    gs_size = vapply(gene_symbol, length, integer(1)),
    de_id = lapply(
      X = ensembl_gene, 
      FUN = intersect, 
      y = dplyr::filter(limma_results, Res.UPF3A_OE_vs_Control != 0)$ensembl_gene_id
      ),
    de_size = vapply(de_id, length, integer(1))
  ) %>% right_join( dplyr::rename(sigRes_H$UPF3A_OE_vs_Control, "gs_name" = fgsea.Geneset)) %>% 
  dplyr::filter(fgsea.padj  < 0.05) %>%
  dplyr::select(gs_name, de_id) %>%
  unnest(de_id) %>%
   mutate(gs_name = str_remove(gs_name, "HALLMARK_")) %>%
  split(f = .$gs_name) %>% 
  lapply(magrittr::extract2, "de_id") %>% 
  fromList() %>% 
  upset(
    nsets = length(.), 
    nintersects = 20,
    order.by = "freq", 
    mb.ratio = c(0.6, 0.4),
    sets.x.label = "Number Of DE Genes"
    )


```
## UPF3A OE, UPF3B KD
```{r}
bind_rows(hm, c2) %>% 
  dplyr::select(gs_name, gene_symbol, ensembl_gene) %>% 
  chop(c(gene_symbol, ensembl_gene)) %>%
  mutate(
    gs_size = vapply(gene_symbol, length, integer(1)),
    de_id = lapply(
      X = ensembl_gene, 
      FUN = intersect, 
      y = dplyr::filter(limma_results, Res.UPF3A_OE_UPF3B_KD_vs_Control != 0)$ensembl_gene_id
      ),
    de_size = vapply(de_id, length, integer(1))
  ) %>% right_join( dplyr::rename(sigRes_H$UPF3A_OE_UPF3B_KD_vs_Contro, "gs_name" = fgsea.Geneset)) %>% 
  dplyr::filter(fgsea.padj  < 0.05) %>%
  dplyr::select(gs_name, de_id) %>%
  unnest(de_id) %>%
   mutate(gs_name = str_remove(gs_name, "HALLMARK_")) %>%
  split(f = .$gs_name) %>% 
  lapply(magrittr::extract2, "de_id") %>% 
  fromList() %>% 
  upset(
    nsets = length(.), 
    nintersects = 20,
    order.by = "freq", 
    mb.ratio = c(0.6, 0.4),
    sets.x.label = "Number Of DE Genes"
    )


```

## Gene sets enriched in all comparisons 

```{r}
enrich_H = list("UPF3B KD" = sigRes_H$UPF3B_KD_vs_Control$fgsea.Geneset, 
                "UPF3A KD" = sigRes_H$UPF3A_KD_vs_Control$fgsea.Geneset, 
                "Double KD" = sigRes_H$DoubleKD_vs_Control$fgsea.Geneset, 
                "UPF3A OE" = sigRes_H$UPF3A_OE_vs_Control$fgsea.Geneset, 
                "UPF3A OE, UPF3B KD" = sigRes_H$UPF3A_OE_UPF3B_KD_vs_Control$fgsea.Geneset)
```


```{r}
upset(fromList(enrich_H),  order.by = "freq")
```
```{r}
inter_data = upset(fromList(enrich_H),  order.by = "freq")
```

```{r fig.height=10, fig.width=10}
sigRes_H_summary = lapply(sigRes_H, function(x) { x %>% dplyr::select(fgsea.Geneset, fgsea.NES) })
sigRes_H_summary = lapply(sigRes_H_summary, function(x)( x %>%  with(structure(fgsea.NES, names=fgsea.Geneset))))

sigRes_H_summary %>% do.call(rbind, .) %>% pheatmap(border_color = "white", cellheight =15, cellwidth = 15, scale = "none")


```

